---
title: | 
       | Lapalican Renormalization 
       | Group (LRG) BA Analysis
pagetitle: "html_booklet"
author: "Begona Bolos & Colin Mclean"
date: "`r Sys.Date()`"
output:
  html_document:
    css: css/cruk.css
    number_sections: true
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
  word_document:
    toc: yes
    toc_depth: '4'
---


```{r setup, include=FALSE}

## load libraries
library(dplyr)
library(tidyr)
library(broom)
library(igraph)
library(poweRlaw)
library(Matrix)
library(expm)
library(ggplot2)
library(latex2exp)
library(Deriv)
library(here)
library(Rbeast)
library(pracma) ## for zeta function


## set directory names
data.dir = here::here("datasets")

knitr::opts_chunk$set(cache=TRUE, echo=TRUE, warning=FALSE)
                      #root.dir=root.dir,
                      #project.dir=project.dir)

## In order to knit a Word doc, using kableExtra library for generating tables,
## 1) clear cache,
## 2) make sure this option is set before loading kableExtra in the Setup.R script.
options(kableExtra.auto_format=TRUE)

## Global options for displaying tables via DT::datatable
#options(DT.options=list(class='cell-border stripe',scrollx=TRUE))

options(DT.fillcontainer=TRUE, 
        DT.options=list(class='cell-border stripe',scrolly="400px", scrollx=TRUE))

## run Setup.R to load libraries globally - by specifying parameter local=knit_global()
#source(, local=knitr::knit_global())
```


```{r get_document_type, warning=FALSE, echo=FALSE, cache=TRUE}
doc.type=NULL 
output <- rmarkdown:::parse_yaml_front_matter(readLines(knitr::current_input()))$output
if( is.list(output) ){
   doc.type=names(output)[1]
   } else {
     doc.type=output[1]
   }
```

```{r, add_header_logo, warning=FALSE, echo=FALSE, cache=FALSE}
css.file=NULL
if(doc.type=="html_document"){
  ## get leading css file used
  css.file = rmarkdown:::parse_yaml_front_matter(readLines(knitr::current_input()))$output$html_document$css[1]
}

if( css.file=="css/cruk.css" ){
  htmltools::img(src =    knitr::image_uri(here::here("css/images/CRUK_SCOTLAND_C_Pos_RGB.png")),
               alt = "CRUK Scotland Centre logo",
               style = "position:absolute; top:0; right:0; padding:10px; height:80px; width:360px;")
}

if( css.file=="css/nhs.css" ){
htmltools::img(src = knitr::image_uri(here::here("css/images/NHS_Lothian.jpg")),
               alt = "NHS Lothian logo",
               style = "position:absolute; top:0; right:0; padding:10px; height:90px; width:90px;")
}

```

<br>
<br>

# Introduction {-}

> Laplacian renomalisation group (LRG) is a dimensional reduction technique inspired from physics. 
> The specific heat of a physical system, ie network, can be used to measure the point at which
> the system transitions from high-dimensional space to a coarse-dimensional space whilst 
> preserving the underlying toplogy of the system.
>
> [1] Villegas, P., Gili, T., Caldarelli, G. et al. Laplacian renormalization group for heterogeneous networks. Nat. Phys. 19, 445–450 (2023). https://doi.org/10.1038/s41567-022-01866-8
>

<br>
<br>

```{r}

## load function required to perform LRG on a network
source(here("core_functions.R"))

```

<br>
<br>

# BA Networks

>
> Probability Distribution of eigenvalues plot for BA network with m=[1,2,3] and 4096 vertices
>

```{r}
ba1 = readRDS(here("_safe_datasets/BA_m1_4096.rds"))
ba1 = ba1[[2]]

ba2 = readRDS(here("_safe_datasets/BA_m2_4096.rds"))
#ba2 = readRDS(here("_safe_datasets/BA_m1_4096_p_eq_2.rds"))
ba2 = ba2[[2]]

ba3 = readRDS(here("_safe_datasets/BA_m3_4096.rds"))
#ba3 = readRDS(here("_safe_datasets/BA_m1_4096_p_eq_3.rds"))
ba3 = ba3[[2]]
```

```{r}

bw  = 2

df1 = bin.eigen(ba1, group="A")
df2 = bin.eigen(ba2, group="B")
df3 = bin.eigen(ba3, group="C")

df1 = smooth.bin.eigen(df1, bandwidth=bw)
df2 = smooth.bin.eigen(df2, bandwidth=bw)
df3 = smooth.bin.eigen(df3, bandwidth=bw)

df  = rbind(df1,df2,df3)

pt1 = ggplot(df, aes(x,y,group=group))+
  geom_line(aes(color=group), linewidth=1)+
  scale_color_manual(
    name="", 
    values=c(A="red", B="green", C="blue"),
    labels=c(A="m=1", B="m=2", C="m=3"))+
   scale_x_log10(
      name = TeX("$\\lambda$"),
      breaks = scales::trans_breaks("log10", function(x) 10^x),
      labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    scale_y_log10(
      name = TeX("P($\\lambda$)"),
      breaks = scales::trans_breaks("log10", function(x) 10^x),
      labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    annotation_logticks(sides="b", outside=TRUE)+
    coord_cartesian(clip = "off")+
    theme_light()+
    theme(legend.position = "right")

```


```{r}
pt1
```


>
> BA scale-free network with 1000 vertices
>

>
> Set the random seed for reproducibility
>


```{r}

## Number of samples
Nits = 1 #500

## Node of nodes
Nv = 1000

## m = [1,2,3]
## default m = 1
m = 1

p=1

## Node names
gn = seq(1,Nv,1)

Lmean      = matrix(0,Nv,Nv)
Esample    = list()
Vsample    = matrix(0,Nv,Nv)
Vinvsample = matrix(0,Nv,Nv)

for( i in 1:Nits ){

## Barabasi Graph
gg=sample_pa(n=Nv, m=m, p=p, directed=FALSE)

## Set node names
V(gg)$name = gn

L = get.L(gg)
E = get.eigen(L, only.values=FALSE, inv.vec=TRUE)

Lmean        = Lmean + as.matrix(L)
Esample[[i]] = E[[1]]
Vsample      = Vsample + E[[2]]
Vinvsample   = Vsample + E[[3]]

}

Lmean   = Lmean/Nits  
Emargin = do.call(rbind, Esample)
Emean   = apply(Emargin, 2, mean)
Esd     = apply(Emargin, 2, sd)
Evec    = Vsample/Nits
Evinv   = Vinvsample/Nits

```

<br>
<br>


>
> Plots
>


```{r}

# Create the first tibble with time intervals from 10^-4 to 10^4 with step 0.1
df1 <- tibble(tau = seq(10^-4, 10^4, by = 0.1))

# Create the second tibble with time intervals from 10^-4 to 1 with step 10^-2
df2 <- tibble(tau = seq(10^-4, 1, by = 10^-2))

# Combine the two tibbles
df  <- bind_rows(df1, df2)

# Remove duplicates and reorder the rows by tau in ascending order
df  <- df %>% 
       distinct() %>%  # Remove duplicates
       arrange(tau)    # Reorder by tau (ascending)
```

> Plot 1-S and find turning points

```{r}

## Plot (1-S)
pt2 = ggplot(df, aes(x=tau))+
     geom_function(fun=S_wrapper, args=list(n=Nv,e=Emean), linewidth=1,
                  aes(color="1-S"))+
    labs(x = "log(tau)", y = "1-S")+
    scale_color_manual(values = c("1-S" = "blue"),
                       breaks = c("1-S"),
                       name = "")+ 
    scale_y_continuous(
    
    ## First axis
    name = "1-S",
    
    ) + 
    scale_x_log10(
        breaks = scales::trans_breaks("log10", function(x) 10^x),
        labels = scales::trans_format("log10", scales::math_format(10^.x))
    ) +
    annotation_logticks(sides="b", outside=TRUE)+
    coord_cartesian(clip = "off")+
    theme_light()+
    theme(legend.position = "right")
```

```{r}
pt2
```

> (1-S) and C curves together

```{r}
## create data for plot
y1      = S_wrapper(e=Emean, t=df$tau, n=Nv)
y2      = dS_dlogt_wrapper(e=Emean, t=df$tau, n=Nv)
not.nan = !is.nan(y2)
y1      = y1[not.nan]
y2      = y2[not.nan]
t.set   = df$tau[not.nan]

```

```{r}

## prepare plot
tmp.y2 = y2

#cut.off=0.1#0.05
cut.off=y2[1]
tmp.y2[tmp.y2<=cut.off]=cut.off

b2        = range(tmp.y2)
b3        = log(seq(b2[1],b2[2], length=4)) 
b3        = exp(b3)
b3        = signif(b3,1)
y2.breaks = unique(c(0,b3))
log.y2    = log(tmp.y2)

a=range(y1)
b=range(log.y2)

## expected analytical specific-heat constant 
#tx = 0.1
#c1 = C1(e=Emean, t=tx)
#C_const = tx*c1*log(Nv)

dat = data.frame(x=t.set, y1=y1, y2=log.y2)

```


```{r}

## create plot
pt3 <- ggplot(dat, aes(x))+
     geom_line(aes(y=y1, color="1-S"), linewidth=1)+
     geom_line(aes(y=(y2-b[1])/diff(b), color="C log(N)"),linewidth=1, alpha=.8)+    
     ##geom_line(aes(y=(y1-a[1])/diff(a), color="1-S"))+
     ##geom_line(aes(y=(y2-b[1])/diff(b), color="Clog(N)"))+ 
     ##geom_hline(yintercept=C_const, linetype="dashed", color="black", linewidth=1)+
    scale_color_manual(values = c("1-S"   = "blue",
                                "C log(N)" = "red"),
                       breaks = c("1-S","C log(N)"),
                       name = "")+ 
     labs(x = "log(tau)")+
     #annotate("rect", xmin=reg.x$tp.min.x_axis[ii], xmax=reg.x$tp.x_axis[ii], 
     #                ymin=0, ymax=1, color="grey", alpha = .3)+
     scale_x_log10()+
     scale_y_continuous(
         name = "1-S",
         labels=~round(.,2),
         breaks=seq(0,1,.25),
         sec.axis=sec_axis(~.*b2[2],
                           name = "log(C log(N))",
                           ##breaks = scales::pretty_breaks(n=5),
                           ####breaks = scales::trans_breaks("log10", function(x) 10^x),
                           ##breaks = scales::breaks_log(4), 
                           labels = scales::label_scientific(digits = 1),
                           breaks = y2.breaks,
                           ##labels = y2.breaks
                           ))+##y2.breaks))+
    annotation_logticks(sides="b", outside=TRUE)+
    coord_cartesian(clip = "off")+
    theme_light()+
    theme(legend.position = "right")
  
```

```{r}
pt3
```

> Plot dC/dt (method 1)

```{r}

## Plot C
pt4 = ggplot(df, aes(x=tau))+
  geom_function(fun=dC_dt_test_wrapper, args=list(n=Nv,e=Emean),
                  linewidth=1, aes(color="dC/dt"))+
    geom_hline(yintercept=0, color="black", linetype="dashed", linewidth=1)+
    labs(x = "log(tau)", y = "")+
    scale_color_manual(values = c("dC/dt" = "red"),
                       breaks = c("dC/dt"),
                       name = "")+ 

    #scale_y_log10(
    scale_y_continuous(
    #limit=c(1e-2,NA),
    ## First axis
    name = "",##dC/dt",
    #breaks = scales::trans_breaks("log10", function(x) 10^x),
    #labels = scales::trans_format("log10", scales::math_format(10^.x))
    
    ) + 
     scale_x_log10(
        breaks = scales::trans_breaks("log10", function(x) 10^x),
        labels = scales::trans_format("log10", scales::math_format(10^.x))
    ) +
    annotation_logticks(sides="b", outside=TRUE)+
    coord_cartesian(clip = "off")+
    theme_light()+
    theme(legend.position = "right")
```

```{r}
pt4
```

> Where does dC/dt cross zero

```{r}
tt.steps = 1e-2
tt      = tibble(tau=seq(1e-4,1e2, tt.steps))
dc      = dC_dt_test_wrapper(n=Nv, t=tt$tau, e=Emean)
dc.diff = diff(sign(dc))
dc.indx = which(abs(dc.diff)==2)
tau.x   = NA
if( length(dc.indx) > 0 ){ tau.x   = tt$tau[dc.indx] }
```

```{r}
print(tau.x)
```

> Plot turning points of dC/dt:

```{r}

tps        = turing.points(pt4, print=0)

```


```{r}
plot(tps$pt)
```

> Calculate graph renormalisation round tau.x

```{r}

##test
t.upper = 5 #tps$tp$tp.x_axis[4] #0.090
t.lower = 0.01 #tps$tp$tp.x_axis[1] #0.060

lrg.scan = scan.time(gg=gg, e=E[[1]], v=E[[2]], vinv=E[[3]], L=L, t.lower=t.lower, t.upper=t.upper,
                      complex=FALSE, method="eigen", order=1, n.steps=150, print=0)

```

```{r}

t.upper = 2 #tps$tp$tp.x_axis[4] #0.090
t.lower = 0.01 #tps$tp$tp.x_axis[1] #0.060


## TEST
lrg.ann = anneal.time(gg=gg, L=L, e=E[[1]], v=E[[2]], vinv=E[[3]], 
                      t.lower=t.lower, t.upper=t.upper, cooling_rate=0.85,
                      complex=FALSE, method="eigen", max_iter=100, dt=1, print=0)
```

```{r}
lrg.ann$df
```

> Best LRG time based on cross-entropy loss:


```{r}
metric="jsd"
x = lrg.scan$df$t
y = lrg.scan$df[,metric]
df.scan = data.frame(x=x, y=y)

pt5 <- ggplot(df.scan, aes(x,y)) +
  geom_line() + 
  ggtitle("Jensson-Shannon Distance between ecdf of network, \n and cdf of fitted lrg network")+
  labs(x="time", y="JSD")+
  geom_point(data=lrg.ann$df, aes(x=new_t,y=Loss.new), color="red")+
  geom_line(data=lrg.ann$df, aes(x=new_t,y=Loss.new), color="red")+
  theme_light() 

```

```{r}
pt5
```

```{r}

## Scan algorithm
indx.best = which.min(lrg.scan$df$jsd)
tau.best  = lrg.scan$df$t[indx.best]

## Annealing algorithm
indx.best1 = nrow(lrg.ann$df)
tau.best1  = lrg.ann$df$new_t[indx.best1]

```

> Plot graph and LRG graph degree distributions

```{r}
lrg.ann$plots[[indx.best1]]
```

```{r}
lrg.ann$plots2[[indx.best1]]
```

> Plot best LRG time on entropy and specific heat plot:

```{r}

## create plot
pt6 <- ggplot(dat, aes(x))+
     geom_line(aes(y=y1, color="1-S"), linewidth=1)+
     geom_line(aes(y=(y2-b[1])/diff(b), color="C log(N)"),linewidth=1, alpha=.8)+    
     ##geom_vline(xintercept=tau.best, linetype="dashed", color="black", linewidth=1)+
    scale_color_manual(values = c("1-S"   = "blue",
                                "C log(N)" = "red"),
                       breaks = c("1-S","C log(N)"),
                       name = "")+ 
     labs(x = "log(tau)")+
     annotate("rect", xmin=t.lower, xmax=t.upper,
                     ymin=0, ymax=1, color="grey", alpha = .3)+
     geom_vline(xintercept=tau.best1, linetype="dashed", color="black", linewidth=1)+
     scale_x_log10()+
     scale_y_continuous(
         name = "1-S",
         labels=~round(.,2),
         breaks=seq(0,1,.25),
         sec.axis=sec_axis(~.*b2[2],
                           name   = "log(C log(N))",
                           labels = scales::label_scientific(digits = 1),
                           breaks = y2.breaks
                           ))+
    annotation_logticks(sides="b", outside=TRUE)+
    coord_cartesian(clip = "off")+
    theme_light()+
    theme(legend.position = "right")
  
```

```{r}
pt6
```


>
> Print size of supernodes
>

```{r}

R = real.LRG(e=E[[1]], v=E[[2]], vinv=E[[3]], L=L, t=tau.best, gg=gg, method="eigen", order=1)
table(R$supernodes)

```